### 2017-06-26 2017/NACO/paper.tex

base a finite number of computed values of the optimized function only in this case .

@@

function \(\varphi(y(x))\), which satisfy the uniform H{\"o}lder condition, i. e.

@@

\begin{equation} \label{step5par}
x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{r|z_{t_j}-z_{t_j-1}|}{m}\right]^N,\; \textrm{ if } 1<t_j<k+1,
\end{equation}
\[
x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
\]
\par
The stopping condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
where the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leq j\leq p.
\end{equation}
If the stopping condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.
\par
The convergence conditions for the developed parallel algorithm are considered
in \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. Thus, in particular, when the convergence conditions are
satisfied, the parallel computations are non-redundant as compared to the serial method
when using up to \(2^N\) processors/cores (\(N\) is the dimensionality of the global optimization problem).

@@

    \item Prior to beginning new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:

@@

    \item After completing any trial for any problem \(\varphi(y^l(x)),1\leq l\leq s\),
    at any point \(x'\in[0,1]\), it is necessary:

@@

      \item to determine all prototypes \(x_i',1\leq i\leq s\) for the point of the

@@

      \(z'=\varphi(y_l(x'))\) to all computational nodes to include the obtained data into

@@

    should check the queue of the received data; if there are any data in the queue,

@@

\section{Results of Numerical Experiments}
\label{sec:experiments}
The numerical experiments were carried out using the Lobachevsky supercomputer at
the State University of Nizhni Novgorod (the operation system – CentOS 6.4, the
supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
E5-2660 2.2 GHz processors with 64 GB RAM. Each processor had 8 cores (i. e. total
16 CPU cores were available at each node). In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
\par
A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores
(total 240 threads). In addition, the supercomputer included the computational nodes equipped
with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
\par
In the performed experiments the optimization problems generated by the GKLS-generator
\cite{gavianoKvasovLeraSergeev2003} were used. This generator allows obtaining the multiextremal optimization
problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc. In order to simulate the computational costs that are inherent in the applied optimization problems, in all experiments computing the objective function values was intentionally complicated by the auxiliary computations not altering the values of the function and the position of its minima.
\par
In all the executed experiments, 100 global optimization problems generated randomly by the GKLS-generator were solved. The results of experiments were averaged over the number of the solved optimization problems.

@@

The numerical comparison was carried out by using the function classes Simple and Hard
of the dimensions 4 and 5. The results of numerical experiments for DIRECT and
DIRECT\textit{l} are taken from \cite{sergeyevKvasov2006}. According to \cite{sergeyevKvasov2006}, the global minimizer \(y^*\)
was considered to be found if the algorithm generated a trial point \(y^k\) inside
a hyperinterval with a vertex \(y^*\) and
\begin{displaymath}
    |y^k(i)-y^*(i)|\le \sqrt[N]{\Delta}(b(i)-a(i)), 1 \le i\le N,
\end{displaymath}
where \(N\) is the problem dimensionality, $a$ and $b$ are the borders of the search domain \(D\)
\footnote{In the case of MAGS, the stop condition was different slightly --- the
global minimizer \(y^*\) was considered to be found, if MAGS generates a trial
point \(y^k\)  in the \(\delta\)-vicinity of the global minimum, i. e. \(||y^k-y^*||\le\delta\).
The size of the vicinity was selected  as \(\delta=\left\|b-a\right\|\sqrt[N]{\Delta}\).}.

@@

\subsection{Parallel Computations Using General Purpose Graphic Processors}
In the first series of experiments, single graphic accelerator was used for solving 100
6-dimensional optimization problems. The reliability parameter \(r=4.5\) was selected.
\par
The averaged execution time using GPU was 10.78 sec whereas the averaged execution
time using all 16 CPU cores of the computational node was 53.8 sec; almost 5-fold improvement was observed.
\par
A larger scale computational experiment was carried out for the global optimization
problems with the dimensions of \(N = 8\) and \(N = 10\): total 12 nodes with 36 graphic
accelerators (three accelerators per the computational node) were employed (total 96768 CUDA-cores were used).
\par
The averaged execution time using the GPUs to solve the 8-dimensional problem was
405.6 sec that was 5.9 times less than the execution time for CPU. The averaged time
of solving the 10-dimensional problem was 2055.8 sec (in this case, the execution time for CPU
was not calculated because of the high complexity level of the required computations).

>>>>

basis of a finite number of computed values of the optimized function only in this case .

@@

function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.

@@

\begin{equation} \label{step5par}
x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{r|z_{t_j}-z_{t_j-1}|}{m}\right]^N,\; \textrm{ if } 1<t_j<k+1,
\end{equation}
\[
x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
\]
\par
The stopping condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
where the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leq j\leq p.
\end{equation}
If the stopping condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.

It should be pointed that in case of PMAGS $p$ trials are performed at each iteration (i.e., $k=np$, where $n$ is the number of executed iteration).

The convergence conditions for the developed parallel algorithm are considered
in \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. Thus, in particular, when the convergence conditions are
satisfied, the parallel computations are non-redundant as compared to the serial method
when using up to \(2^N\) processors/cores (\(N\) is the dimensionality of the global optimization problem).

@@

    \item Prior to beginning each new trial at some point \(x'\in [0,1]\) for a problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:

@@

    \item After completing some trial for a problem \(\varphi(y^l(x)),1\leq l\leq s\),
    at some point \(x'\in[0,1]\), it is necessary:

@@

      \item to determine all prototypes \(x_i',1\leq i\leq s,\) for the point $x'\in[0,1]$ of the

@@

      \(z'=\varphi(y^l(x'))\) to all computational nodes to include the obtained data into

@@

    should check the queue of the received data; if there are some data in the queue,

@@

\section{Results of Numerical Experiments}
\label{sec:experiments}
The numerical experiments were carried out using the Lobachevsky supercomputer at
the State University of Nizhni Novgorod (the operation system – CentOS 6.4, the
supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
E5-2660 2.2 GHz processors with 64 GB RAM. Each processor had 8 cores (i. e. total
16 CPU cores were available at each node). In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
\par
A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores. In addition, the supercomputer included the computational nodes equipped
with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (in total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
\par
In the performed experiments the optimization problems generated by the GKLS-generator
\cite{gavianoKvasovLeraSergeev2003} were used. This generator allows obtaining the multiextremal optimization
problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc. In order to simulate the computational costs that are inherent in the applied optimization problems, in all experiments computing the objective function values was intentionally complicated by the auxiliary computations not altering the values of the function and the position of its minima.
\par
In all the executed experiments, 100 global optimization problems generated randomly by the GKLS-generator were solved. The results of experiments were averaged over the number of the solved optimization problems.

@@

The numerical comparison was carried out by using the function classes Simple and Hard
of the dimensions 4 and 5. The results of numerical experiments for DIRECT and
DIRECT\textit{l} are taken from \cite{sergeyevKvasov2006}. According to \cite{sergeyevKvasov2006}, the global minimizer \(y^*\)
was considered to be found if the algorithm generated a trial point \(y^k\) inside
a hyperinterval with a vertex \(y^*\) and
\begin{displaymath}
    |y^k_i-y^*_i|\le \sqrt[N]{\Delta}(b_i-a_i), 1 \le i\le N,
\end{displaymath}
where \(N\) is the problem dimensionality, $a$ and $b$ are the borders of the search domain \(D\)
\footnote{In the case of MAGS, the stopping condition was different slightly --- the
global minimizer \(y^*\) was considered to be found, if MAGS generates a trial
point \(y^k\)  in the \(\delta\)-vicinity of the global minimum, i. e. \(||y^k-y^*||\le\delta\).
The size of the vicinity was selected  as \(\delta=\left\|b-a\right\|\sqrt[N]{\Delta}\).}.

@@

\subsection{Parallel Computations Using General Purpose Graphic Processors}
In the first series of experiments, single graphic accelerator was used for solving 100
6-dimensional optimization problems. The reliability parameter \(r=4.5\) was selected.
\par
The averaged execution time using GPU was 10.78 sec whereas the averaged execution
time using all 16 CPU cores of the computational node was 53.8 sec; almost 5-fold improvement was observed.
\par
A larger scale computational experiment was carried out for the global optimization
problems with the dimensions of \(N = 8\) and \(N = 10\): total 12 nodes with 36 graphic
accelerators (three accelerators per the computational node) were employed (in total 96768 CUDA-cores were used).
\par
The averaged execution time using the GPUs to solve the 8-dimensional problem was
405.6 sec that was 5.9 times less than the execution time for CPU. The averaged time
of solving the 10-dimensional problem was 2055.8 sec (in this case, the execution time for CPU
was not calculated because of the high complexity level of the required computations).
